---
title: "`ggplot`"
author: "Xiaoyue Ma"
date: "5/10/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

devtools::install_github("thomasp85/patchwork")

library(tidyverse)
library(ggridges)
library(patchwork)
```

## Create the weather data

```{r load_data, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

## create a ggplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

Alternative way of making this plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

print plot

```{r}
scatterplot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

scatterplot


```

## Color

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5)
```

## Why do `aes` positions matter?

first
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```

vs.
```{r}
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE)
```

vs.
```{r}
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE, aes(color = name))
```

## facet

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(~name)
 ```
```

this is fine, but not interesting

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE, size = 2)
 ```
```

## some new stuff

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(se = FALSE, size = 2)
```

## 2D density

```{r}
#install packeges("hexbin")
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_bin2d()
 ```
```

 ## more kinds of plots!


```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(~name)
```

## density plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)
 ```
```

##box plot

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

## violin plot
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

## ridge plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```


## making new plots

start with an old plot.

```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
 ```
```

## add labels

```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    tittle = "Temperature plot", 
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15(C)", "-5(still very cold)", "20(nice weather")
  ) +
  scale_y_continuous(
    trans = "sqrt"
  )
  
 ```
```


## Colors next

```{r}
ggp_base = 
  weather_df %>% 
  ggplot(aes(x =tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    tittle = "Temperature plot", 
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_base
 ```
```

## themes

```{r}
ggp_base +
  theme(legend.position = "bottom")
 ```
```

```{r}
ggp_base +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## more than one dataset

```{r}
central_park = 
   weather_df %>%
   filter(name == "CentralPark_NY")
 
Waikiki = 
   weather_df %>%
   filter(name == "Waikiki_HA")

ggplot(data = Waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp)) +
  geom_line(data = central_park)
 


```

## brief aside about colors

```{r}
Waikiki %>% 
  ggplot(aes(x = date, y = tmax)) + 
  geom_point(color = "blue")
```


## multi-panel plots

```{r}
gg_scatter = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

gg_density = 
  weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_density()

gg_box = 
  weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
  

gg_scatter + (gg_density / gg_box)


```

## data manipulation

```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_relevel(name, "Waikiki_HA", "CentralPark_NY")
  ) %>% 
  ggplot(aes(x = name, y = tmin, color = name)) +
  geom_boxplot()
```

## reorder instead of relevel

```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_reorder(name, tmax)
  ) %>% 
  ggplot(aes(x = name, y = tmin, color = name)) +
  geom_boxplot()
```

```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_relevel(name, "Waterhole_WA", "Waikiki_HA", "CentralPark_NY")
  ) %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()

```

## restructure then plot

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x = temperature, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(~name) +
  theme(legend.position = "bottom")
```

litters and pups

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))

fas_data = left_join(pup_data, litter_data, by = "litter_number")
```

```{r}
fas_data %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_day"
  ) %>%
    drop_na(dose, day_of_tx, pn_day) %>%
    mutate(
      outcome = factor(outcome),
      outcome = fct_reorder(outcome, pn_day)
    ) %>% 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
  facet_grid(day_of_tx ~ outcome)

```


```{r}
fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, day_of_tx, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```






